Testing Gravity Using Type Ia Supernovae Discovered by Next Generation Wide Field Imaging Surveys

### 2019-03-11 doc/article.tex
affiliates done

M.~White\footnotemark[1],\textsuperscript{,}\footnotemark[20],

@@

\textbf{Endorsers:}
 \linebreak

@@

%\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
% $^{1}$ \ED \\
%$^{2}$ \Durham \\
%$^{3}$ \SISSA \\
%$^{4}$ \IFPU \\
%$^{5}$ \INFN \\
%$^{6}$ \CFT \\
%$^{7}$ \LBL \\
%$^{8}$ \Pitt \\
%$^{9}$ \IFT \\
%$^{10}$ \UAM \\
%$^{11}$ \KIPAC \\
%$^{12}$ \Stanford \\
%$^{13}$ \SLAC \\
%$^{14}$ \Yale \\
%$^{15}$ \UR \\
%$^{16}$ \dunlap \\
%$^{17}$ \daa \\
%$^{18}$ \Umich \\
%$^{19}$ \JHU \\
%$^{20}$ \Port \\
%$^{21}$ \UNIPD \\
%$^{22}$ \SMU \\
%$^{23}$ \IUCAA \\
%$^{24}$ \UGTO \\
%$^{25}$ \FNAL \\
%$^{26}$ \RomaS \\
%$^{27}$ \INFNRM \\
%$^{28}$ \USC \\
%$^{29}$ \STSCI \\
%$^{30}$ \KSU \\
%$^{31}$ \StonyBrook \\
%$^{32}$ \KASSI \\
%$^{33}$ \SHAO\\
%$^{34}$ \BNL \\
%$^{35}$ \Syracuse \\
%$^{36}$ \UCB \\
%$^{37}$ \HarvardPhys \\
%$^{38}$ \NAOC \\
%$^{39}$ \UCAS \\
%

@@

Peculiar velocity surveys have already been  used to measure $fD$ (also referred to as $f\sigma_8$), though not to a level where  gravity models can be precisely distinguished.
 \cite{2017MNRAS.471..839A} use 6dFGS peculiar velocities using  Fundamental Plane distances of elliptical galaxies to estimate absolute magnitudes
 with
 $\sim 0.43$~mag  precision, yielding a 15\% uncertainty in $fD$ at $z\approx 0$.
The upcoming 
TAIPAN survey \cite{2017PASA...34...47D} will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
and the WALLABY+WNSHS surveys \cite{2008ExA....22..151J} will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
of galaxies with densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
$z=0-0.1$ covering 75\% of the sky.
These surveys combined are projected to have 3\% uncertainties in $fD$ \cite{2017MNRAS.464.2517H}.
For reference, DESI projects a 10\% precision of $fD$ at $z \approx 0.3$  by looking 
for signatures (Redshift Space Distortions; RSD) expected from galaxies infalling toward mass overdensities.
Relative to galaxies with  Fundamental Plane or Tully-Fisher distances, 
SN~Ia host galaxies currently have significantly lower number density but have better per-object peculiar velocity precision.
Existing SN~Ia samples
have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
\cite{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
SNe~Ia discovered by ZTF \cite{2019PASP..131a8002B} over the next several years will provide first probative measures of $fD$ at $z<0.1$.
%Measurement of the velocity field using LSST-discovered
%SNe~Ia has been quantified by \cite{2011PhRvD..83d3004B,2017JCAP...01..060O}

Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter spectral model yields  absolute magnitude
dispersion $\sigma_M \gtrsim 0.12$~mag.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
Recent studies indicate that with the right data, SN absolute
magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \cite[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}. 
Though not yet
established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
that is observed using current calibrations.  At this precision the intrinsic velocity dispersion  at $z=0.028$ is  of $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
measure a peculiar velocity with $S/N \sim 1$.
 If corrections of all SNe~Ia are not possible, the use of SN~Ia subclasses is an option though at the expense of reducing the
numbers of velocity probes.
Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
  will increase the number of identified  $z<0.3$ Type~Ia supernovae from the hundreds to the
hundreds of thousands; over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
 for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$.
  This sample has comparable
 number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.  With similar densities,
 the (two) ten-year SN~Ia survey will have
 a (6) 29$\times$ reduction in shot-noise, $\sigma^2_M/n$, relative to the Fundamental Plane survey of TAIPAN.

@@

Survey performance is examined in more detail by considering how $\sigma_\gamma$ in the cross-correlation analysis
changes with respect to the 
survey parameters  $\Omega$, $z_{\rm max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  Though not directly a survey parameter, we also examine changes with respect to our fiducial choice of  $k_{\rm max}$. \newline
%
%The dependence on the uncertainty in the growth index on the survey parameters is as follows:
%
{\it Solid Angle $\Omega$:}
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
%
{\it Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{\rm max}=0.2$.
The left panel of Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution, the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$. 
\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{src/dvardz.png}
\includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
\caption{Left: $|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survy with limiting  depth $z_{\rm max}=0.2$. 
Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
 $z_{\rm max}$ for two- (dashed) and ten-year (solid) surveys.
\label{dvar:fig}}
\end{figure}\newline
{\it Redshift Depth $z_{\rm max}$:}
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{\rm max} \sim 0.2$ as seen in Figure~\ref{var:fig}. \newline
{\it Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
An increased survey duration accumulates more supernovae, decreasing shot noise and  increasing the 
precision in $\gamma$ for all the surveys considered.
The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution (i.e.\ 
shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration, as shown
n the right-panel plot of Figure~\ref{dvar:fig} that shows $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ as a function of
 $z_{\rm max}$ for two- and ten-year surveys.
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has a similar relationship with $\sigma_\gamma$ as shown
in the left-panel plot of $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$  of Figure~\ref{dvar:fig}.\newline
{\it Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.  Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{\rm max} = 0.0050$ at $k_{\rm max}=0.1h$\,Mpc$^{-1}$,
 which  is survey-independent.

>>>>

M.~White\footnotemark[1]\textsuperscript{,}\footnotemark[17],

@@

\input{/Users/akim/project/AstroInstitutions/institutionAliases.tex}
\textbf{Endorsers:}
Behzad Ansarinejad
\footnotemark[22],
\footnotetext[22]{Department of Physics, Lower Mountjoy, South Rd, Durham DH1 3LE, United Kingdom} 
Robert Armstrong\footnotemark[23],
\footnotetext[23]{Lawrence Livermore National Laboratory, Livermore, CA, 94550} 
Jacobo Asorey\footnotemark[18],
Carlo Baccigalupi\footnotemark[24]\textsuperscript{,}\footnotemark[25]\textsuperscript{,}\footnotemark[26],
\footnotetext[24]{SISSA - International School for Advanced Studies, Via Bonomea 265, 34136 Trieste, Italy} 
\footnotetext[25]{IFPU - Institute for Fundamental Physics of the Universe, Via Beirut 2, 34014 Trieste, Italy} 
\footnotetext[26]{INFN - National Institute for Nuclear Physics, Via Valerio 2, I-34127 Trieste, Italy} 
Maciej Bilicki\footnotemark[27],
\footnotetext[27]{Center for Theoretical Physics, Polish Academy of Sciences, al. Lotnik\'{o}w 32/46, 02-668, Warsaw, Poland} 
Julian Borrill\footnotemark[16], 
Elizabeth Buckley-Geer\footnotemark[28],
\footnotetext[28]{Fermi National Accelerator Laboratory, Batavia, IL 60510} 
Kelly A. Douglass\footnotemark[4],
Cora Dvorkin\footnotemark[29],
\footnotetext[29]{Department of Physics, Harvard University, Cambridge, MA 02138, USA} 
Simon Foreman\footnotemark[30],
\footnotetext[30]{Canadian Institute for Theoretical Astrophysics, University of Toronto, Toronto, ON M5S 3H8, Canada} 
Llu\'is Galbany\footnotemark[21],
Juan Garc\'ia-Bellido\footnotemark[31]\textsuperscript{,}\footnotemark[32],
\footnotetext[31]{Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain} 
\footnotetext[32]{Universidad Aut\'{o}noma de Madrid, 28049, Madrid, Spain} 
Martina Gerbino\footnotemark[33],
\footnotetext[33]{HEP Division, Argonne National Laboratory, Lemont, IL 60439, USA} 
Mandeep S.S. Gill\footnotemark[34]\textsuperscript{,}\footnotemark[35]\textsuperscript{,}\footnotemark[36],
\footnotetext[34]{Kavli Institute for Particle Astrophysics and Cosmology, Stanford 94305} 
\footnotetext[35]{Stanford University, Stanford, CA 94305} 
\footnotetext[36]{SLAC National Accelerator Laboratory, Menlo Park, CA 94025} 
Larry Gladney\footnotemark[37],
\footnotetext[37]{Department of Physics, Yale University, New Haven, CT 06520} 
Saurabh W.~Jha\footnotemark[38],
\footnotetext[38]{Department of Physics and Astronomy, Rutgers, the State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854, USA} 
Johann Cohen-Tanugi\footnotemark[39],
\footnotetext[39]{Laboratoire Univers et Particules de Montpellier, Univ. Montpellier and CNRS, 34090 Montpellier, France} 
D.\ O.\ Jones\footnotemark[40],
\footnotetext[40]{University of California at Santa Cruz, Santa Cruz, CA 95064} 
Marc Kamionkowski\footnotemark[41],
\footnotetext[41]{Johns Hopkins University, Baltimore, MD 21218} 
Ryan E. Keeley\footnotemark[18],
Robert Kehoe\footnotemark[42],
\footnotetext[42]{Southern Methodist University, Dallas, TX 75275} 
Savvas M. Koushiappas\footnotemark[43],
\footnotetext[43]{Brown University, Providence, RI 02912} 
Ely D.~Kovetz\footnotemark[44],
\footnotetext[44]{Department of Physics, Ben-Gurion University, Be'er Sheva 84105, Israel} 
Kazuya Koyama\footnotemark[45],
\footnotetext[45]{Institute of Cosmology \& Gravitation, University of Portsmouth, Dennis Sciama Building, Burnaby Road, Portsmouth PO1 3FX, UK} 
Benjamin L'Huillier\footnotemark[18],
Ofer Lahav\footnotemark[46],
\footnotetext[46]{University College London, WC1E 6BT London, United Kingdom} 
Chien-Hsiu Lee\footnotemark[47],
\footnotetext[47]{National Optical Astronomy Observatory, 950 N. Cherry Ave., Tucson, AZ 85719 USA} 
Michele Liguori\footnotemark[48],
\footnotetext[48]{Dipartimento di Fisica e Astronomia ``G. Galilei'',Universit\`a degli Studi di Padova, via Marzolo 8, I-35131, Padova, Italy} 
Axel de la Macorra\footnotemark[49],
\footnotetext[49]{IFUNAM - Instituto de F\'{i}sica, Universidad Nacional Aut\'onoma de M\'exico, 04510 CDMX, M\'exico} 
Joel Meyers\footnotemark[42],
Surhud More\footnotemark[50],
\footnotetext[50]{The Inter-University Centre for Astronomy and Astrophysics, Pune, 411007, India} 
Jeffrey A. Newman\footnotemark[21],
Gustavo Niz\footnotemark[51],
\footnotetext[51]{Divisi\'{o}n de Ciencias e Ingenier\'{i}as, Universidad de Guanajuato, Le\'{o}n 37150, M\'{e}xico} 
Antonella Palmese\footnotemark[28],
Francesco Piacentini\footnotemark[52]\textsuperscript{,}\footnotemark[53],
\footnotetext[52]{Dipartimento di Fisica, Universit\`{a} La Sapienza, P. le A. Moro 2, Roma, Italy} 
\footnotetext[53]{Istituto Nazionale di Fisica Nucleare, Sezione di Roma, 00185 Roma, Italy} 
Steven Rodney\footnotemark[54],
\footnotetext[54]{The University of South Carolina, Columbia, SC 29208} 
Benjamin Rose\footnotemark[55],
\footnotetext[55]{Space Telescope Science Institute, Baltimore, MD 21218} 
Masao Sako\footnotemark[56],
\footnotetext[56]{Department of Physics and Astronomy, University of Pennsylvania, Philadelphia, Pennsylvania 19104, USA} 
Lado Samushia\footnotemark[57],
\footnotetext[57]{Kansas State University, Manhattan, KS 66506} 
Neelima Sehgal\footnotemark[58],
\footnotetext[58]{Stony Brook University, Stony Brook, NY 11794} 
Leonardo Senatore\footnotemark[34],
Sara Simon\footnotemark[13],
Stephanie Escoffier\footnotemark[59],
\footnotetext[59]{Aix Marseille Univ, CNRS/IN2P3, CPPM, Marseille, France} 
Scott Watson\footnotemark[60],
\footnotetext[60]{Syracuse University, Syracuse, NY 13244} 
Weishuang Xu\footnotemark[29],
Gong-Bo Zhao\footnotemark[61]\textsuperscript{,}\footnotemark[45],
\footnotetext[61]{National Astronomical Observatories, Chinese Academy of Sciences, PR China} 
Yi Zheng\footnotemark[62]
\footnotetext[62]{School of Physics, Korea Institute for Advanced Study, 85 Hoegiro, Dongdaemun-gu, Seoul 130-722, Korea}

@@

% $^{1}$ \Durham \\
%$^{2}$ \LLNL \\
%$^{3}$ \KASSI \\
%$^{4}$ \SISSA \\
%$^{5}$ \IFPU \\
%$^{6}$ \INFN \\
%$^{7}$ \CFT \\
%$^{8}$ \LBL \\
%$^{9}$ \FNAL \\
%$^{10}$ \Queensland \\
%$^{11}$ \UR \\
%$^{12}$ \HarvardPhys \\
%$^{13}$ \CITA \\
%$^{14}$ \Pitt \\
%$^{15}$ \IFT \\
%$^{16}$ \UAM \\
%$^{17}$ \ANLHEP \\
%$^{18}$ \KIPAC \\
%$^{19}$ \Stanford \\
%$^{20}$ \SLAC \\
%$^{21}$ \Yale \\
%$^{22}$ \dunlap \\
%$^{23}$ \daa \\
%$^{24}$ \UMich \\
%$^{25}$ \Rutgers \\
%$^{26}$ \LUPM \\
%$^{27}$ \UCSC \\
%$^{28}$ \JHU \\
%$^{29}$ \SMU \\
%$^{30}$ \Brown \\
%$^{31}$ \BenGurion \\
%$^{32}$ \Port \\
%$^{33}$ \UCL \\
%$^{34}$ \NOAO \\
%$^{35}$ \UNIPD \\
%$^{36}$ \IFUNAM \\
%$^{37}$ \LPC \\
%$^{38}$ \IUCAA \\
%$^{39}$ \UGTO \\
%$^{40}$ \RomaS \\
%$^{41}$ \INFNRM \\
%$^{42}$ \USC \\
%$^{43}$ \STSCI \\
%$^{44}$ \UPenn \\
%$^{45}$ \KSU \\
%$^{46}$ \StonyBrook \\
%$^{47}$ \BNL \\
%$^{48}$ \CPPM \\
%$^{49}$ \Syracuse \\
%$^{50}$ \UCB \\
%$^{51}$ \NAOC \\
%$^{52}$ \KIAS \\

@@

Peculiar velocity surveys have already been  used to measure $fD$ (also referred to as $f\sigma_8$), though not to a level where  gravity models can be precisely distinguished.
 \cite{2017MNRAS.471..839A} use 6dFGS peculiar velocities using  Fundamental Plane distances of elliptical galaxies to estimate absolute magnitudes
 with
 $\sim 0.43$~mag  precision, yielding a 15\% uncertainty in $fD$ at $z\approx 0$.
The upcoming 
TAIPAN survey \cite{2017PASA...34...47D} will obtain Fundamental Plane galaxies with densities of $n_g \sim 10^{-3}h^3$\,Mpc$^{-3}$,
and the WALLABY+WNSHS surveys \cite{2008ExA....22..151J} will obtain Tully-Fisher distances (based on the $\sim 0.48$~mag calibration of absolute magnitude based on the  HI 21cm line width)
of galaxies with densities $n_g \sim 2\times 10^{-2} - 10^{-4} h^3$\,Mpc$^{-3}$ from
$z=0-0.1$ covering 75\% of the sky.
These surveys combined are projected to have 3\% uncertainties in $fD$ \cite{2017MNRAS.464.2517H}.
For reference, DESI projects a 10\% precision of $fD$ at $z \approx 0.3$  by looking 
for signatures (Redshift Space Distortions; RSD) expected from galaxies infalling toward mass overdensities.
Relative to galaxies with  Fundamental Plane or Tully-Fisher distances, 
SN~Ia host galaxies currently have significantly lower number density but have better per-object peculiar velocity precision.
Existing SN~Ia samples
have been used to test and ultimately find spatial correlations in peculiar velocities that may be attributed to the growth of structure
\cite{PhysRevLett.99.081301,2008MNRAS.389L..47A,2014MNRAS.444.3926J,2015JCAP...12..033H, 2017JCAP...05..015H}.
SNe~Ia discovered by ASAS-SN, ATLAS, and ZTF \cite{2014ApJ...788...48S,2018PASP..130f4505T,2019PASP..131a8002B} over the next several years will provide first probative measures of $fD$ at $z<0.1$.
%Measurement of the velocity field using LSST-discovered
%SNe~Ia has been quantified by \cite{2011PhRvD..83d3004B,2017JCAP...01..060O}

Two advances in the upcoming decade will make SN~Ia peculiar velocities more powerful.
First, the precision of SN~Ia distances can be improved.  The commonly-used empirical 2-parameter spectral model yields  absolute magnitude
dispersion $\sigma_M \gtrsim 0.12$~mag.  However, SNe transmit more information than just the light-curve shape and single color used in current SN models.
Recent studies indicate that with the right data, SN absolute
magnitudes can be calibrated to $\sigma_M \lesssim 0.08$ mag \cite[see e.g.][]{2012MNRAS.425.1007B, 2015ApJ...815...58F}. 
Though not yet
established, it is anticipated that such a reduction in intrinsic dispersion comes with a reduction in the magnitude bias correlated with host-galaxy properties
that is observed using current calibrations.  At this precision the intrinsic velocity dispersion  at $z=0.028$ is  $300$~km\,s$^{-1}$, i.e.\ a single SN~Ia  is of such quality as to
measure a peculiar velocity with $S/N \sim 1$.
 If corrections of all SNe~Ia are not possible, the use of SN~Ia subclasses is an option though at the expense of reducing the
numbers of velocity probes.
Secondly,  in the upcoming decade cadenced wide-field imaging surveys such as ZTF and LSST
  will increase the number of identified  $z<0.3$ Type~Ia supernovae from the hundreds to the
hundreds of thousands; over the course of 10-years, LSST will find $\sim150,000$ $z<0.2$,  $\sim520,000$ $z<0.3$ SNe~Ia
 for which good light curves can be measured, corresponding to a  number density of $n \sim 5\times 10^{-4}h^3$\,Mpc$^{-3}$.
  This sample has comparable
 number density and more galaxies at deeper redshifts than projected by WALLABY and TAIPAN.  With similar densities,
 the (two) ten-year SN~Ia survey will have
 a (6) 29$\times$ reduction in shot-noise, $\sigma^2_M/n$, relative to the Fundamental Plane survey of TAIPAN.

@@

Survey performance is examined in more detail by considering how $\sigma_\gamma$ in the cross-correlation analysis
changes with respect to the 
survey parameters  $\Omega$, $z_{\rm max}$, $t$, and $\sigma_M$, and also with respect to differential redshift bins
within a given survey.  Though not directly a survey parameter, we also examine changes with respect to our fiducial choice of  $k_{\rm max}$. \newline
%
%The dependence on the uncertainty in the growth index on the survey parameters is as follows:
%
{\it Solid Angle $\Omega$:}
The Fisher Matrix $F$ is  proportional to the survey solid angle $\Omega$ so $\sigma_\gamma \propto \Omega^{-1/2}$.
%
{\it Differential Redshift Bin $z$:}  Certain redshifts constrain $\gamma$ more strongly than others.  If at a given moment of a survey we had
a set of SNe~Ia from which to choose, it turns out the one with the lowest redshift would be preferred.
This is demonstrated to be the case at the end of both 2- and 10-year surveys with $z_{\rm max}=0.2$.
The left panel of Figure~\ref{dvar:fig} shows
 $|\partial \sigma_\gamma / \partial z|$, which
for both surveys monotonically decreases from $z=0.01$ out to $z=0.2$.  If we had
to sculpt the distribution, the preference would be to cut out the highest redshift bins
resulting in a decreased $z_{\rm max}$.   The optimal redshift distribution is thus the unsculpted SN-discovery distribution truncated by $z_{\rm max}$. 
\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{src/dvardz.png}
\includegraphics[width=0.49\textwidth]{src/dvardxxx.png}
\caption{Left: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial z|$  after two and ten years for a survey with limiting  depth $z_{\rm max}=0.2$. 
Right: $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ (red) and $\sigma_\gamma^{-1}\partial \sigma_\gamma / \partial \sigma_M$ (black) each as a function of
 $z_{\rm max}$ for two- (dashed) and ten-year (solid) surveys.
\label{dvar:fig}}
\end{figure}\newline
{\it Redshift Depth $z_{\rm max}$:}
Increasing the survey redshift depth increases the $\gamma$ precision.  The differential improvement in $\sigma_\gamma$ plateaus at
$z_{\rm max} \sim 0.2$ as seen in Figure~\ref{var:fig}. \newline
{\it Survey duration $t$; Intrinsic Magnitude Dispersion $\sigma_M$:}
An increased survey duration accumulates more supernovae, decreasing shot noise and  increasing the 
precision in $\gamma$ for all the surveys considered.
The surveys we consider have varying relative contributions of sample variance and shot noise: those that have a larger shot-noise contribution (i.e.\ 
shorter surveys and those with higher $z_{\rm max}$) benefit more from extending the survey duration.
% in the right-panel plot of Figure~\ref{dvar:fig}.
Like survey duration, intrinsic magnitude dispersion is related to survey performance through the shot noise and thus
has a similar relationship with $\sigma_\gamma$; the effect of duration and magnitude dispersion are shown in the right-panel plot of Figure~\ref{dvar:fig} as  $\sigma_\gamma^{-1}|\partial \sigma_\gamma / \partial \ln{t}|$ 
and $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial \sigma_M$ 
as a function of
 $z_{\rm max}$ for two- and ten-year surveys.\newline
{\it Minimum length scale, maximum wavenumber $k_{\rm max}$:} There is a minimum length scale at which density and velocity distributions are reliably predicted from theory.  Changes in this scale engender  fractional changes in the $\gamma$ precision as  $\sigma_\gamma^{-1}\partial \sigma_\gamma/\partial k_{\rm max} = 0.0050$ at $k_{\rm max}=0.1h$\,Mpc$^{-1}$,
 which  is survey-independent.
